---
title: "Uncertainty Estimation in the AONR and EONR implementing a quadratic plateau Bayesian model"
author: Francisco Palmero
output:
  html_document:
    df_print: paged
---

Bayesian inference is implemented to estimate the uncertainty on the "optimal" N rates (AONR, EONR). 

The fitted model for maize grain yield as a response of N rates was a quadratic-plateau with block random effect in the intercept:

Likelihood
$$y_{ij} \sim normal(\mu_{ij}, \sigma_\epsilon),$$ where


\[
\mu_{ij}=
\begin{cases}
\beta_0 + \tau_j + \beta_1x_i - \beta_2x_i^2, &\text{if } x_i\leq AONR\\
\beta_0 + \tau_j + \beta_1(AONR) - \beta_2(AONR)^2, &\text{o.w.}
\end{cases}
\],

where $$AONR = \frac{\beta_1}{2\beta_2}$$


The parameter models:
$$\beta_0 \sim gamma(\alpha_{\beta_{0_k}}, \beta_{\beta_{0_k}}),$$
$$\beta_1 \sim gamma(\alpha_{\beta_{1_k}}, \beta_{\beta_{1_k}}),$$
$$\beta_2 \sim gamma(\alpha_{\beta_{2_k}}, \beta_{\beta_{2_k}}),$$
$$\tau_j \sim normal(0,\sigma_b)$$
$$\sigma_b \sim uniform(0,1000)$$
$$\sigma_\epsilon \sim gamma(\alpha_{\sigma_{\epsilon_k}}, \beta_{\sigma_{\epsilon_k}})$$


By fitting this model the data, we can obtain the agronomic and the economic optimum N rates, AONR and EONR, respectively.
$$AONR = \frac{\beta_1}{2\beta_2},$$ and

$$EONR = \frac{\beta_1-PR}{2\beta_2}$$.

To include the variability in maize and N fertilizer prices:
-> For fertilizers, we retrieved the average price of anhydrous ammonia (82-0-0) and urea (46-0-0) at each April or March.
-> For maize, we considered the future price at each April. 
-> These data were retrieved from USDA-ERS and USDA-NASS for the period 2000 - 2021.  
-> For each iteration of the MCMC, the PR was sampled from $PR \sim gamma(15.2, 2.85)$ and included in the calculus of the EONR. 



# Calling the requiered packages
```{r libraries, warning=FALSE, message=FALSE}
library("rjags")
```


The example you will see below contains the hyperparameters for the models fitted for the sites inside the state of Iowa

# 1) Specify the model
Blocks Random Effects on Intercept
```{r}
model_b0r <- "model{

# Loop through the 4 blocks (K) to create a vector of random deviations.
for (k in 1:K){ # product expression over k
  b0_b[k] ~ dnorm(0, 1/sigma_block^2)
} # end of k product

for (i in 1:N) { # product expression over i
  
  # In b0_b[block[i]], 'block' should be passed as data. 
  # So, in b0_b[block[i]], block[i]: Retrieves the block indicator for the current observation i
  # and b0_b[block[i]]: Uses the block indicator to access the corresponding random effect intercept from the vector b0_b
  
  # Deterministic model
  mu[i] <- ifelse(x[i] <= (b1/(2*b2)), 
  b0 + b0_b[block[i]] + b1*x[i] - b2*(x[i]^2),
  b0 + b0_b[block[i]] + b1*(b1/(2*b2)) - b2*((b1/(2*b2))^2))
  
  # Likelihood
  y[i] ~ dnorm(mu[i], 1/sigma^2)

} # end of i product

# Priors
b0 ~ dgamma(10.89,0.00155)
b1 ~ dgamma(3.26,0.0467)
b2 ~ dgamma(1.5,10.00)

# Priors on variances
sigma_block ~ dunif(0,5000)
sigma ~ dgamma(3.75,0.0025)

# Uncertainty around price ratio
PR ~ dgamma(15.2, 2.85)
# Obtain the AONR
AONR <- b1/(2*b2)
# Obtain the EONR while accounting for the uncertainity in the price ratio
EONR <- (b1-PR)/(2*b2)

}"

```


# 2) Create a function to fit the model
```{r}
bayes_model <- function(x,y,block,
                        parameters = c("b0","b1","b2",
                                       "b0_b","sigma_block","sigma", 
                                       "AONR", "EONR"),
                        n.chains = 4, 
                        n.iter = 60000, 
                        n.adapt = 30000,
                        thin = 10,
                        seed = 5897
                        ){
  
  data <- list(y = y,
               x = x,
               N = length(y),
               block = block,
               K = length(unique(block)))
  
  set.seed(seed)
  jags_model <- jags.model(textConnection(model_b0r), 
                           data = data,
                           n.adapt = n.adapt,
                           n.chains = n.chains)
  
  post_samples <- coda.samples(jags_model, 
                               variable.names = parameters, 
                               n.iter = n.iter,
                               thin = thin)
  
  return(post_samples)
  
}

```


# 3) Fit the model to the data
```{r}
model_saved <- bayes_model(x=df$N_rate, 
            y=df$Yield,
            block=df$Block)
```






